Network Origins of Social Distinction

### 2025-07-26 manuscript.tex
Updates from Overleaf

Thus, distinction is zero for all nodes in the circle graph. The same reasoning applies to the clique graph in which nodes have the same constant status scores as in the circle graph. The difference is that in this case the $\kappa$ component of Equation~\ref{eq:distinction} is obtained from the status scores of each actor in the node-deleted subgraph excluding the focal node, which is just a clique of size $N-1$ where $N$ is the number of nodes in the original clique. Just like in the circle graph case, however, both the $s$ and $\kappa$ components are constant across nodes, which means that the status scores cancel out by division and the constraint scores by subtraction, leaving everyone with a scaled distinction score of zero.\footnote{Because of the local nature of the $\kappa$ component in Equation~\ref{eq:distinction}, in a regular $(n, k)$ graph of order $n$ with prescribed degree $k$ across nodes, we should not expect scaled distinction to be exactly zero, since the focal nodes previous neighbors will have different status scores than the other nodes in the graph in the focal-node-deleted subgraph. However, we should not expect their status to be wildly different, since their degree is now only $k-1$. Instead, we should expect scaled distinction to fluctuate \textit{around} the zero value across nodes. Nevertheless, these should cancel out across nodes, and therefore we should also expect the \textit{average} distinction across nodes in a regular $(n, k)$ graph to be zero.}

@@

%Triadic closure 3 versions
Psychology
Polarity, valenced ties 
Status distinction

>>>>

Thus, the distinction is zero for all nodes in the circle graph. The same reasoning applies to the clique graph in which nodes have the same constant status scores as in the circle graph. The difference is that in this case the $\kappa$ component of Equation~\ref{eq:distinction} is obtained from the status scores of each actor in the node-deleted subgraph, excluding the focal node, which is just a clique of size $N-1$ where $N$ is the number of nodes in the original clique. Just like in the circle graph case, however, both the $s$ and $\kappa$ components are constant across nodes, which means that the status scores cancel out by division and the constraint scores by subtraction, leaving everyone with a scaled distinction score of zero.\footnote{Because of the local nature of the $\kappa$ component in Equation~\ref{eq:distinction}, in a regular $(n, k)$ graph of order $n$ with prescribed degree $k$ across nodes, we should not expect scaled distinction to be exactly zero, since the focal nodes previous neighbors will have different status scores than the other nodes in the graph in the focal-node-deleted subgraph. However, we should not expect their status to be wildly different, since their degree is now only $k-1$. Instead, we should expect scaled distinction to fluctuate \textit{around} the zero value across nodes. Nevertheless, these should cancel out across nodes, and therefore we should also expect the \textit{average} distinction across nodes in a regular $(n, k)$ graph to be zero.}

@@

%Triadic closure 3 versions
\textit Triadic closure is an observed regularity in which a connection is likely to form between two actors who, although not originally connected, are first connected to a common alter (Simmel 1908). In essence, friends of friends have an increased likelihood of also becoming friends. The causes of this regularity, and the reasons for its occasional failure, are largely based on consideration of valence. Positive or negative affect drive the resulting triadic relations. The original consideration was for the cognitive balance of the individual, or the psychological congruity of affect by an individual for two objects that they see as causally connected. As such, the valenced relationship of any three actors are brought into combined consideration because any two are considered affectively combined through their joint combination as "the parts of a causal unit (Heider 1946:107)."

\citep{huang2018will}
\citep{yoon2013exchange}
\citep{mollenhorst2011shared} 
\citep{feld1981focused}
\citep{feld1982social}

Psychology
Polarity, valenced ties 
Status distinction
